	   \begin{frame}\frametitle{Inference Systems - Soundness (Recap)}

\begin{itemize}
\item
  \alert{An inference is sound} if 
  the conclusion of this inference is a logical 
  consequence of its premises.

\item
  \alert{An inference system
  is sound} if every inference rule in this system is sound.
\end{itemize}

\bigskip

\vs<2->{
\Blue{$\M{\BRis}$ is sound.}

\bigskip

Consequence of soundness: let $S$ be a set of clauses. If
$\emptyclause$ can be derived from $S$ in $\BRis$, then
$S$ is \alert{unsatisfiable}.

\bigskip\medskip

\Blue{Lecture 2 - Exercise recap : infinite number of $\BRis$
  derivations of~$\emptyclause$}
}


                           \end{frame}

%---------------------------------------------------------------------

	   \begin{frame}
\frametitle{Can this be used for checking (un)satisfiability (Recap)}


\begin{enumerate}
  \item What happens when the empty clause \alert{cannot be derived}
  from $\M{S}$?
  \item \alert{How} can one search for possible derivations of the empty clause?
\end{enumerate}

                           \end{frame}

%---------------------------------------------------------------------

	   \begin{frame}
\frametitle{Can this be used for checking (un)satisfiability (Recap)}

\begin{enumerate}\item
\alert{Completeness.}

\begin{quotation}\noindent\Blue{%
Let $\M{S}$ be an unsatisfiable set of clauses. Then there exists a 
derivation of $\M{\emptyclause}$ from $\M{S}$ in $\M{\BRis}$.}
\end{quotation}

\item \vs<2->{
We have to formalize \alert{search for derivations.}}
\end{enumerate}

\bigskip 
\vs<3->{
  We introduced \alert{well-behaved selection functions} for selecting
  literals in clauses and applying inferences only over selected literals.}


\bigskip

\vs<4->{Binary resolution $\BRis$ with selection is \alert{complete for every
  well-behaved selection function}}.

                           \end{frame}
